maze-dataset: Maze Generation with Algorithmic Variety and Representational Flexibility

### 2025-04-26 docs/paper/paper.tex
move acknowledgements

Each
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
is constructed from a set of several
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
objects, each of which specifies how different token regions or other
elements of the stringification are produced.

@@

The tokenizer architecture is purposefully designed such that adding and
testing a wide variety of new tokenization algorithms is fast and
minimizes disturbances to functioning code. This is enabled by the
modular architecture and the automatic inclusion of any new tokenizers
in integration tests. To create a new variety of tokenizer, developers
forking the library may simply create their own
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
subclass and implement the abstract methods. If the behavior change is
sufficiently small, simply adding a parameter to an existing
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
subclass and updating its implementation will suffice.

The breadth of tokenizers is also easily scaled in the opposite
direction. Due to the exponential scaling of parameter combinations,
adding a small number of new features can significantly slow certain
procedures which rely on constructing all possible tokenizers, such as
integration tests. If any existing subclass contains features which
aren't needed, a developer tool decorator
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization/modular/element_base.html\#mark_as_unsupported}{\texttt{@mark\_as\_unsupported}}
is provided which can be applied to the unneeded
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
subclasses to prune those features and compact the available space of
tokenizers.

@@

\begin{figure}
\hypertarget{fig:benchmarks}{%
\centering
\includegraphics[width=0.95\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
\caption{Plot of maze generation time. Generation time scales
exponentially with maze size for all algorithms. Generation time per
maze does not depend on the number of mazes being generated, and there
is minimal overhead to initializing the generation process for a small
dataset. Wilson's algorithm is notably less efficient than others and
has high variance. Note that values are averaged across all parameter
sets for that algorithm. More information can be found on the
\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
page}.}\label{fig:benchmarks}
}
\end{figure}

@@

This work was partially funded by National Science Foundation awards
DMS-2110745 and DMS-2309810. We are also grateful to LTFF and FAR Labs
for hosting authors MII, AFS, and TR for a residency visit, and to
various members of FAR's technical staff for their advice.

This work was partially supported by AI Safety Camp and AI Safety
Support, which also brought many of the authors together. We would like
to thank our former collaborators at AI Safety Camp and other users and
contributors to the \texttt{maze-dataset} package: Benji Berczi,
Guillaume Corlouer, William Edwards, Leon Eshuijs, Chris Mathwin, Lucia
Quirke, Can Rager, Adrians Skapars, Rusheb Shah, Johannes Treutlein, and
Dan Valentine.

We thank the Mines Optimization and Deep Learning group (MODL) for
fruitful discussions. We also thank Michael Rosenberg for recommending
the usage of Finite State Transducers for storing tokenizer validation
information.

>>>>

Each
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
is constructed from a set of several
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
objects, each of which specifies how different token regions or other
elements of the stringification are produced.

@@

The tokenizer architecture is purposefully designed such that adding and
testing a wide variety of new tokenization algorithms is fast and
minimizes disturbances to functioning code. This is enabled by the
modular architecture and the automatic inclusion of any new tokenizers
in integration tests. To create a new variety of tokenizer, developers
forking the library may simply create their own
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
subclass and implement the abstract methods. If the behavior change is
sufficiently small, simply adding a parameter to an existing
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
subclass and updating its implementation will suffice.

The breadth of tokenizers is also easily scaled in the opposite
direction. Due to the exponential scaling of parameter combinations,
adding a small number of new features can significantly slow certain
procedures which rely on constructing all possible tokenizers, such as
integration tests. If any existing subclass contains features which
aren't needed, a developer tool decorator
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization/modular/element_base.html\#mark_as_unsupported}{\texttt{@mark\_as\_unsupported}}
is provided which can be applied to the unneeded
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#_TokenizerElement}{\texttt{\_TokenizerElement}}
subclasses to prune those features and compact the available space of
tokenizers.

@@

\begin{figure}
\hypertarget{fig:benchmarks}{%
\centering
\includegraphics[width=0.9\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
\caption{Plot of maze generation time. Generation time scales
exponentially with maze size for all algorithms. Generation time per
maze does not depend on the number of mazes being generated, and there
is minimal overhead to initializing the generation process for a small
dataset. Wilson's algorithm is notably less efficient than others and
has high variance. Note that values are averaged across all parameter
sets for that algorithm. More information can be found on the
\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
page}.}\label{fig:benchmarks}
}
\end{figure}

@@

\input{figures/tex/acknowledgements.tex}
